---
title: "BH explanation / visualization"
author: "Nikos Ignatiadis"
date: "`r doc_date()`"
package: "`r pkg_ver('IHWpaper')`"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{"BH explanation / visualization"}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

Below we just generate the necessary plot to explain how BH works.

```{r}
library("ggplot2")
library("dplyr")
library("wesanderson")
library("grid")
library("gridExtra")
library("IHW")
```

# B.Sc. thesis plot (low $\pi_0$)

Plot as in B.Sc. thesis with very low $\pi_0$. (Using this so it can be clearly demonstrated that the BH threshold is an intermediate threshold between the Bonferroni threshold and the uncorrected one, also such $\pi_0$ allows to show all p-values in the second plot and still observe the interesting behaviour.)

```{r}
simpleSimulation <- function(m,m1,betaA,betaB){
    pvalue <- runif(m)
    H <- rep(0,m)
    alternatives <- sample(1:m,m1)
    
    pvalue[alternatives] <- rbeta(m1,betaA,betaB)
    H[alternatives] <-1
 
    simDf <- data.frame(pvalue = pvalue, group=runif(m), filterstat = runif(m), H=H)
    return(simDf)
}

set.seed(1)
sim <- simpleSimulation(1000, 700, 0.3, 8)
sim$rank <- rank(sim$pvalue)

histogram_plot <- ggplot(sim, aes(x=pvalue)) + 
                    geom_histogram(binwidth=0.1, fill = wes_palette("Chevalier1")[4]) + 
                    xlab("p-value") + 
                    theme_bw()
                        
bh_threshold <- get_bh_threshold(sim$pvalue, .1)

bh_plot <- ggplot(sim, aes(x=rank, y=pvalue)) +
  geom_step(col=wes_palette("Chevalier1")[4]) + 
  ylim(c(0,0.2)) +
  geom_abline(intercept=0, slope= 0.1/1000, col = wes_palette("Chevalier1")[2]) +
  geom_hline(yintercept=bh_threshold, linetype=2) + 
  annotate("text",x=250, y=0.065, label="BH testing") + 
  geom_hline(yintercept = 0.1, linetype=2) + 
  annotate("text",x=250, y=0.11, label="uncorrected testing") + 
  geom_hline(yintercept = 0.1/1000, linetype=2) +
  annotate("text",x=850, y=0.1/1000+0.01, label="Bonferroni testing") + 
  ylab("p-value") + xlab("rank of p-value") +
  theme_bw() + scale_colour_manual(values=wes_palette("Chevalier1")[c(3,4)]) 

```

```{r fig.width=11, fig.height=5}
grid.arrange(histogram_plot, bh_plot, nrow=1)
```

```{r eval=FALSE}
pdf(file="bh_explanation.pdf", width=11, height=5)
grid.arrange(histogram_plot, bh_plot, nrow=1)
dev.off()
```

# Bioc presentation plot (higher $\pi_0$)

For ddhw presentation, remake above plot with higher $\pi_0$.
```{r}
set.seed(1)
sim <- simpleSimulation(10000, 2000, 0.3, 8)
sim$rank <- rank(sim$pvalue)

histogram_plot <- ggplot(sim, aes(x=pvalue)) + 
                    geom_histogram(binwidth=0.1, fill = wes_palette("Chevalier1")[4]) + 
                    xlab("p-value") + 
                    theme_bw(14)
                        

bh_threshold <- get_bh_threshold(sim$pvalue, .1)

bh_plot <- ggplot(sim, aes(x=rank, y=pvalue)) +
  geom_step(col=wes_palette("Chevalier1")[4], size=1.2) + 
  scale_x_continuous(limits=c(0,2000),expand = c(0, 0))+
  scale_y_continuous(limit=c(0,0.06), expand=c(0,0)) +
  geom_abline(intercept=0, slope= 0.1/10000, col = wes_palette("Chevalier1")[2], size=1.2) +
  annotate("text",x=500, y=1.3*bh_threshold, label="BH rejection threshold") + 
  geom_hline(yintercept=bh_threshold, linetype=2, size=1.2) + 
  ylab("p-value") + xlab("rank of p-value") +
  theme_bw() + scale_colour_manual(values=wes_palette("Chevalier1")[c(3,4)])
```
```{r fig.width=11, fig.height=5}
grid.arrange(histogram_plot, bh_plot, nrow=1)
```


```{r eval=FALSE}
pdf(file="bh_explanation_high_pi0.pdf", width=11, height=5)
grid.arrange(histogram_plot, bh_plot, nrow=1)
dev.off()
```