---
title: "Explaining tdr"
author: "Nikos Ignatiadis"
date: "October 12, 2015"
output: html_document
---

In this vignette, one of the supplemental figures is generated, which clarifies some concepts regarding the tdr. In particular, the tdr at a fixed p-value is illustrated, the tdr as a function of p-values and also tdr(P_i) as a random variable, which in general will not be a test statistic. It's only a test statistic when we know pi0 and the alternative density, and in such a case it really provides higher discriminatory power than the p-values.


```{r warning=F, message=F}
library("ggplot2")
library("dplyr")
library("tidyr")
library("gridExtra")
library("cowplot")
```

Helper functions to generate simulated data and also evaluate tdr and other quantities on them.
```{r warning=F}
falt1 <- function(t){
    1-pnorm(qnorm(1-t)-mu1)
  }

  lfdr1 <- function(t){
    pi10/(pi10 + (1-pi10)*falt1_density(t))
  }
  falt1_density <- function(t){
      dnorm(qnorm(1-t)-mu1)/dnorm(qnorm(1-t))
  }

#http://stackoverflow.com/questions/10081479/solving-for-the-inverse-of-a-function-in-r
inverse = function (f, lower = -100, upper = 100) {
   function (y) uniroot((function (x) f(x) - y), lower = lower, upper = upper)[1]$roo
}

sim_pars <- function(eff_size, pi0){
  distrib <- function(t) pi0*t + (1-pi0)*(1-pnorm(qnorm(1-t)-eff_size))
  alt_density <-   function(t) dnorm(qnorm(1-t)-eff_size)/dnorm(qnorm(1-t))
  density <- function(t) pi0 + (1-pi0)*alt_density(t)
  tdr <- function(t)  1- pi0/density(t)
  inverse_f <- function(y) uniroot((function (x) density(x) - y), lower = 0.001, upper = 0.999)[1]$root
  tdr_density <- function(t) pi0*(inverse_f(pi0/(1-t)+0.001) - inverse_f(pi0/(1-t)-0.001))/0.002
  return(list(pi0=pi0,distrib=distrib, alt_density=alt_density,density=density, tdr=tdr, inverse_f=inverse_f,
            tdr_density=tdr_density  ))
}

````

Helper function to generate plots for different parameter values:
```{r warning=F}
plot_tdr <- function(x,pi0,xthreshold=0.05){
  eff_size <- x
  pars <- sim_pars(x,pi0)
  df <- data.frame(t=seq(0.001,1, length=1000))
  df <- mutate(df, f1 = pars$alt_density(t), f=pars$density(t), tdr = pars$tdr(t))
  plot_pv <- ggplot(df, aes(x=t, y=f)) + 
                     geom_line()  + 
                     ylim(c(0,4)) + 
                     geom_vline(xintercept=xthreshold, col="blue") + 
                     geom_hline(yintercept=pars$pi0,col="red") +
                     #geom_hline(yintercept=pars$density(xthreshold), linetype=2)+
                     ylab("density")+
                     xlab("p-value")+
                     scale_x_continuous(expand = c(0, 0)) 
  
  plot_pv_vs_tdr <- ggplot(df, aes(x=t, y= tdr)) + 
                    geom_line() +
                    ylab("tdr")+
                    xlab("p-value")+
                    scale_x_continuous(expand = c(0, 0)) 
  
  # now also generate empirical histogram plots
  a = sim_pars(eff_size,pi0)
  m=100000
  H <- rbinom(m, 1, 1-pi0)
  X <- rnorm(m) + eff_size*H
  PV <- 1 - pnorm(X)
  TDR <- a$tdr(PV)
  
  sim_df <- data.frame(pvalue=PV, tdr=TDR)
  
  hist_pv <- ggplot(sim_df, aes(x=pvalue))
  grid.arrange(plot_pv, plot_pv_vs_tdr, ncol=1)
  
  pv_hist <- ggplot(sim_df, aes(x=pvalue)) + geom_histogram(aes(y = ..density..),binwidth=0.036, colour = "darkgreen", fill = "white") + xlab("p-value")
  tdr_hist <- ggplot(sim_df, aes(x=tdr)) + geom_histogram(aes(y = ..density..),binwidth=0.036, colour = "darkgreen", fill = "white") + xlab("tdr")
  list(plot_pv=plot_pv,  plot_pv_vs_tdr=plot_pv_vs_tdr, pv_hist=pv_hist, tdr_hist=tdr_hist)
}
```
Generate all plots:
```{r fig.keep="none", warning=F}
myplots <- c(plot_tdr(3.5, 0.6), plot_tdr(3.5,0.85), plot_tdr(1.5,0.6))
myplots <- myplots[c(1,5,9) + rep(0:3,each=3)]
myplots <- c(myplots, list(labels = letters[1:12]), list(ncol=3, align="hv"))
```

Let's put all plots together:
```{r warning=F, fig.width=10, fig.height=15}
do.call(plot_grid,myplots)
```

```{r eval=F}
pdf("pvalue_vs_tdr_explanation.pdf", width=10, height=15)
do.call(plot_grid,myplots)
dev.off()
```