---
title: "Weights vs Independent Filtering"
author: "Nikos Ignatiadis"
date: "October 12, 2015"
output: html_document
---

In this vignette we show that sometimes the weights can behave very similarly to an independent filtering threshold:

```{r warning=F, message=F}
library("DDHW")
library("ddhwPaper")
library("ggplot2")
library("dplyr")
library("wesanderson")
```

First we generate the simulated data and then use (data-driven greedy) independent filtering. We convert the binary threshold cut-off into weights.

```{r}
sim <- du_ttest_sim(80000, 0.9, 2, seed=1)
ddhf_res <- ddhf(sim$pvalue, sim$filterstat, .1)
ws <- ifelse(sim$filterst <= ddhf_res$cutoff_value, 0, 1)
ws <- ws/sum(ws)*length(sim$pvalue)
```

The induced weight function has total regularization equal to the following:
```{r}
total_regularization_lambda <- max(ws)-min(ws)
total_regularization_lambda
```

We now apply DDHW based on the above total regularization penalization parameter. This makes the comparison more directly applicable.

```{r warning=F}
ddhw_res <- ddhw(sim$pvalue, sim$filterstat, .1, lambdas = total_regularization_lambda, nfolds=1L,  nbins=20)
```

The plot below compares the binary threshold with the DDHW threshold, which is a bit smoother:

```{r fig.width=5, fig.height=5}
df <- rbind( data.frame(covariate = sim$filterstat, weight= ws, method="filtering"),
             data.frame(covariate = sim$filterstat, weight= weights(ddhw_res, levels_only=F), 
                        method=paste0("DDHW; \nlambda=",format(total_regularization_lambda,digits=2))))
weights_filter_plot <- ggplot(df, aes(x=covariate, y=weight, col=method)) + geom_step(size=1.65)+
                      scale_colour_manual(values=wes_palette("Cavalcanti")[c(1,2)]) +
                      theme_classic(16)
weights_filter_plot
```

```{r, eval=FALSE}
pdf("smoothed_threshold.pdf", width=5, height=5)
weights_filter_plot
dev.off()
```

But weights can be even more flexible if we allow the total regularization parameter to vary. Here we use 5fold CV along an equally spaced grid of lambdas to find the best total variation regularization using DDHW:

```{r message=F, warning=F}
ddhw_res2 <- ddhw(sim$pvalue, sim$filterstat, .1, lambdas=seq(0,10,length=20), nfolds=1L, nfolds_internal = 5L, nbins=20, quiet=T)
```

The returned regularization parameter is the following:

```{r message=F, warning=F}
ddhw_res2@regularization_term
```
Now we plot everything:
```{r fig.width=5, fig.height=5}
df <- rbind(df,
            data.frame(covariate = sim$filterstat, weight= weights(ddhw_res2, levels_only=F),  
                       method=paste0("DDHW; \nlambda=",
                                     format(ddhw_res2@regularization_term ,digits=2))))
            
weights_filter_plot <- ggplot(df, aes(x=covariate, y=weight, col=method)) + geom_step(size=1.65)+
                      scale_colour_manual(values=wes_palette("Cavalcanti")[c(1,2,3)]) +
                      theme_classic(16)

weights_filter_plot
```

Notice that all 3 types of weights still fullfill the necessary budget condition:

```{r}
group_by(df, method) %>% summarize(mean_weight = mean(weight))
```
